---
title: "A Descriptive Analysis of Leni Robredo's Career as a Legislator"
author: "Abe Ceasar Perez"
date: "November 1, 2021"
fontsize: 8pt
output:
  beamer_presentation: default
  ioslides_presentation: default
header-includes:
  - \usepackage{booktabs}
---

## Foreward

I have created this presentation as a way for Filipinos to further understand the milestones of one of the presidential candidates that I support. With the amount of misinformation in social media, I feel compelled to use my expertise as a data analyst/statistician to synthesize the right information to further understand the advocacies of my selected candidate. I fervently hope that with this information, the readers will get a better understanding of what my preferred candidate possesses and why she should be the preferred candidate among other contenders.

---

## Introduction

As the filling of COC approached the set deadline, ongoing vice-president Leni Robredo has finally filed her candidacy to run for the presidency of the country. This has sparked positive news across the social media platforms, particularly from supporters and artists who have been longing to hear the announcement. Critics, however, continue to denounce and attack the move made by her excellency as they continued to spread attacks and misinformation in an attempt to dismantle her campaign. Despite the figures that the supporters have continued to disseminate regarding the VP's contribution during the pandemic, critics are still clamoring for any hard evidence as to what she can bring into the table.



The aim of this presentation is to understand the legislative career of the candidate as a congresswoman. This also aims to:

- Determine the most significant contributions of Robredo as a legislator
- Describe the milestones of Robredo using tables and charts
- Infer the main platforms and focuses of Robredo based on the visualizations

---

## Findings

The table below summarizes the legislative track record of Leni Robredo. From 2013 to 2016, she has authored or co-authored a total of 122 legislations - 31 of which are primarily authored while the remaining 91 are either co-authored or sponsored. A total of 14 laws were enacted under her term.

```{r table1, echo = FALSE, warning=FALSE}
setwd("C:\\Users\\Abe\\Desktop\\Elections 2022\\Campaign for Leni")

#Loading the packages

suppressPackageStartupMessages(library(rvest))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggwordcloud))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))

#Load the saved HTML file named [Last name, first name].html

link <- "Robredo, Leni.html"
last_name <- str_match(link, "(.+), ")[,2]
first_name <- str_match(link, ", (.+).html")[,2]
page = read_html(link)

#Extract each text from the html file and convert it into a vector of strings

extract_1 = page %>% html_nodes("p:nth-child(5)") %>% html_text()
extract_2 = page %>% html_nodes("p:nth-child(4)") %>% html_text()
extract_3 = page %>% html_nodes("p:nth-child(6)") %>% html_text()
extract_4 = page %>% html_nodes(".panel-body") %>% html_text()
extract_5 = page %>% html_nodes(".text-muted-size") %>% html_text()
combined_extracts <- list(extract_1, extract_2, extract_3, extract_4, extract_5)

#Further clean each vectorized strings and assign it with the appropriate naming column
#And create the final dataframe by joining all of the cleaned string of vectors into columns

create_dataset <- function(pulled_data){
    dates_filed <- str_match(pulled_data[[2]][1:160], "Date Filed: ([0-9-]+)")[,2]
    house_bill_no <- pulled_data[[5]]
    full_title <- str_remove_all(pulled_data[[1]][1:160], "Full Title: ")
    short_title <- str_match(pulled_data[[3]], "Short Title:(.+)Principal")[,2]
    republic_act_title <- str_match(pulled_data[[4]], "Republic Act Title:(.+)Bill Status")[,2]
    significance <- str_match(pulled_data[[4]], "Significance: (.+)Date Filed")[,2]
    category <- str_match(pulled_data[[3]][1:160], "Referral[s]?: ([A-Z ]+)")[,2]
    author <- str_match(pulled_data[[3]][1:160], "Author/s: (.+)Date")[,2]
    status <- str_replace_all(str_match(pulled_data[[4]][1:160], "Bill Status: (.+)")[,2],"Mother Bill Status:",", Mother Bill Status:")
    republic_act_no <- str_match(status, "Republic Act (.+) enacted")[,2]
    
    
    combined <- data.frame(cbind(dates_filed, house_bill_no, full_title, short_title, republic_act_no, republic_act_title, significance, category, author, status))
    final <- combined %>% mutate(dates_filed = ymd(dates_filed), 
                                          legislation = if_else(str_detect(house_bill_no, "HB"), "bill", "resolution"),
                                          authorship = if_else(str_detect(author, "GERONA-ROBREDO"),"primary author","co-author"), 
                                          enacted_into_law = if_else(str_detect(status, "(?i)enacted"), "yes","no"),
                                          date_of_enactment = ymd(str_match(status, "enacted on (.+)")[,2])) 
    
    final$category[c(131,130,81)] <- c("AGRARIAN REFORM", "RULES", "LOCAL GOVERNMENT")
    final <- final %>% filter(legislation == "bill")
    return(final)
}
legislations <- create_dataset(combined_extracts)

main_stats <- data.frame(length(unique(year(legislations$dates_filed)))-1,
                         nrow(legislations),
                         sum(legislations$authorship=="primary author"),
                         sum(legislations$authorship=="co-author"),
                         14)
colnames(main_stats) <- c("Years of Service","Total no. of legislations filed/supported", "Number of legislations authored",
                          "Number of legislations co-authored or sponsored","Number of bills enacted into law")
df1 <- t(main_stats)
kbl(df1, booktabs=T) %>% kable_styling(font_size = 8)

```

---

## Perfomance in the 16th Congress

```{r chart_16, echo = FALSE, warning=FALSE}
final_df <- read.csv("16th_congress_data.csv")
final_df %>% ggplot(aes(x=total_bills_filed)) +geom_histogram(fill="pink",color="pink4",bins=25) +
  geom_histogram(data=subset(final_df,total_bills_filed < 122),
                 fill="pink3",color="pink4", bins=25)+annotate(geom="text", x=435, y=64, label="Robredo filed more bills than nearly", size=6.5)+annotate(geom="text", x=435, y=64, label=paste("\n\n",round(100*(sum(final_df$total_bills_filed < 122)/length(final_df$total_bills_filed)),2),"% of all the members in the 16th Congress",sep=""), size=6, fontface="bold")+labs(title="How did Robredo perform against her colleagues in the 16th Congress?",subtitle ="A histogram showing the distribution of total bills filed per congressperson",caption = "Source: Congress.gov.ph/legis",x="Total bills filed",y="")+
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 1.8),
                                                                                                                                                      plot.subtitle = element_text(size=15, face = "italic", hjust =0.5),
                                                                                                                                                      plot.caption = element_text(size=15, face="italic",vjust=0.1),
                                                                                                                                                      axis.text = element_text(size=15), axis.title = element_text(size=15),
                                                                                                                                                      panel.background = element_rect(fill = "white", colour = "white"))
```

---

## Perfomance in the 16th Congress

```{r chart_17, echo = FALSE, warning=FALSE}
final_df %>% ggplot(aes(x=no_of_days)) +geom_histogram(fill="pink",color="pink4",bins=25) +
  geom_histogram(data=subset(final_df,no_of_days < 114),
                 fill="pink3",color="pink4", bins=25)+annotate(geom="text", x=175, y=30, label="Robredo was more active than", size=6.5)+annotate(geom="text", x=175, y=30, label=paste("\n\n",round(100*(sum(final_df$no_of_days < 114)/length(final_df$no_of_days)),2),"% of all the members of the 16th Congress", sep=""), size=6, fontface="bold") + labs(title="How active was Robredo against her colleagues in the 16th Congress?",subtitle ="A histogram showing the total number of days in congress per congressperson",caption = "Source: Congress.gov.ph/legis",x="Total days active",y="")+
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 1.8),
                                                                                                                                                      plot.subtitle = element_text(size=15, face = "italic", hjust =0.5),
                                                                                                                                                      plot.caption = element_text(size=15, face="italic",vjust=0.1),
                                                                                                                                                      axis.text = element_text(size=15), axis.title = element_text(size=15),
                                                                                                                                                      panel.background = element_rect(fill = "white", colour = "white"))
```

---

## Visualizing Robredo's legislations and sponsorships

```{r plot1, echo = FALSE, warning=FALSE}

create_word_freq <- function(data){
  bill_words <- data.frame(table(unlist(str_match_all(data, "([A-Z]+)"))))
  bill_words <- bill_words %>% arrange(-Freq) %>% rename(Word = Var1)
  filtered_words <- c("OTHER", "REPUBLIC", "THEREFOR", "PURPOSE", "PURPOSES", "OTHERWISE", "KNOWN",
                      "WITH", "PHILIPPINES", "PHILIPPINE", "LEAST", "FROM", "HOUSE", "REPRESENTATIVES",
                      "THROUGH", "EVERY", "YEAR", "THEIR", "THAT", "THEREOF", "THREE", "EACH", "INTO","RESOLUTION",
                      "CODE", "AMENDED","JOINT","SECTION","ENTITLED","BASED")
  bill_words_three_or_more <- bill_words %>% filter(str_length(Word) > 3 & !(Word %in% filtered_words) & !(str_detect(Word, "(?i)ing")))
  bill_words_three_or_more$Word <- str_to_lower(bill_words_three_or_more$Word)
  bill_words_three_or_more <- bill_words_three_or_more %>% mutate(Freq = round(Freq, 3), angle = 90 * sample(c(0, 1), n(), replace = TRUE, prob = c(60, 40)))
  return(bill_words_three_or_more)
}
bill_words <- create_word_freq(legislations$full_title)

ggplot(data = bill_words[c(1:100),], aes(label = Word, size = Freq, angle = angle, color = as.character(Freq))) + geom_text_wordcloud(max_steps = 1,grid_size = 1, eccentricity = .9) + scale_size_area(max_size = 18) + labs(title=paste("What were the key areas that ",last_name," focused on her legislations?",sep=""), 
                                      subtitle="A wordcloud showing the most frequent words found for each bill title",
                                      caption = "Source: Congress.gov.ph/legis") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
                                                                                         plot.subtitle = element_text(size=15, face = "italic", hjust = 0.5),
                                                                                         plot.caption = element_text(size=15, face="italic"),
                                                                                         panel.background = element_rect(fill = "white"))

```

---

## Visualizing Robredo's legislations and sponsorships

```{r plot2, echo=FALSE}
#Creating a bar chart to identify which committees do the candidates' bills usually belong

bill_categories <- legislations %>% group_by(category) %>% summarize(no_of_bills = n()) %>% arrange(-no_of_bills)
bill_categories[c(1:10),] %>% ggplot(aes(x=no_of_bills,y=reorder(category,no_of_bills))) + geom_col(fill="pink3") + labs(title=paste("Which categories did ",last_name,"'s bills usally fall under?", sep=""), subtitle = "A bar chart showing the distribution of bills per primary committee referred to",
                                                                                                                                 caption = "Source: Congress.gov.ph/legis",x="", y="") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 1.8),
                                                                                                                                                      plot.subtitle = element_text(size=15, face = "italic", hjust =2.2),
                                                                                                                                                      plot.caption = element_text(size=15, face="italic",vjust=0.1),
                                                                                                                                                      axis.text = element_text(size=15),
                                                                                                                                                      panel.background = element_rect(fill = "white", colour = "white"))
```

---

## Visualizing Robredo's legislations and sponsorships

```{r plot3, echo=FALSE}
#Creating a pie chart to determine how the candidate's authorship is distributed for all bills
table_authorship <- legislations %>% group_by(authorship) %>% summarize(bills = n(), percent = round(100*(n()/nrow(legislations)),2)) %>% arrange(-percent)
table_authorship %>% ggplot(aes(x="", y = bills, fill=authorship)) + geom_bar(stat="identity",width = 1,colour="grey") + coord_polar(theta="y",direction=-1) +
  theme_void() + geom_text(aes(y = c(80,20), label = paste(percent,"%",sep="")), color = "black", size=8) +
  scale_fill_manual(values=c("pink2","pink3")) + guides(fill=guide_legend(title="Bill Authorship"))+labs(title=paste("How much of ",last_name,"'s legislations were primarily authored?",sep=""), 
                                                                                                        subtitle="A pie chart showing the distribution of bills by authorship ",
                                                                                                        caption = "Source: Congress.gov.ph/legis") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0),
                                                                                                                                                           plot.subtitle = element_text(size=15, face = "italic", hjust = 2),
                                                                                                                                                           plot.caption = element_text(size=15, face="italic", hjust=1.8,vjust=3.5),
                                                                                                                                                           legend.title = element_text(size=15, face="bold"),
                                                                                                                                                           legend.text = element_text(size=15),
                                                                                                                                                           panel.background = element_rect(fill = "white", colour="white"))
```

---

## Visualizing Robredo's legislations and sponsorships

```{r plot5, echo=FALSE}
#Creating a pie chart to determine the significance of all bills authored/co-authored by the candidate

table_significance <- legislations %>% filter(!(is.na(significance)))%>% group_by(significance) %>% summarize(bills = n(), percent = round(100*(n()/sum(!(is.na(legislations$significance)))),2)) %>% arrange(-percent)
table_significance %>% ggplot(aes(x="", y = bills, fill=significance)) + geom_bar(stat="identity",width = 1,colour="grey") + coord_polar(theta="y", start=0) + theme_void() + geom_text(aes(y = c(55,115), label = paste(percent,"%",sep="")), color = "black", size=6) + scale_fill_manual(values=c("pink2","pink3")) + guides(fill=guide_legend(title="Significance")) + labs(title=paste("How significant are ",last_name,"'s authored/co-authored bills?",sep=""), 
                                                                                                        subtitle="A pie chart showing the distribution of bills by significance",
                                                                                                        caption = "Source: Congress.gov.ph/legis") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.1),
                                                                                                                                                           plot.subtitle = element_text(size=15, face = "italic", hjust = 1.5),
                                                                                                                                                           plot.caption = element_text(size=15, face="italic", hjust=1.6),
                                                                                                                                                           legend.title = element_text(size=15, face="bold"),
                                                                                                                                                           legend.text = element_text(size=15),
                                                                                                                                                           panel.background = element_rect(fill = "white", colour="white"))
```

---

## Enacted Bills

```{r table4, echo = FALSE, warning=FALSE}
options(knitr.kable.NA = '')
#Enacted bills
enacted_bills <- legislations %>% filter(enacted_into_law == "yes") %>% select(dates_filed, house_bill_no, full_title, short_title, authorship, significance, category, republic_act_no, republic_act_title, date_of_enactment) %>% arrange(date_of_enactment)

df4 <- enacted_bills[,c("date_of_enactment","republic_act_no","short_title")]
colnames(df4) <- c("Date of Enactment","Republic Act No." ,"Title")
df4 <- subset(df4, !(is.na(Title)))
df4$Title[c(2,4,5,11,12)] <- c("Extending The Corporate Life Of The Philippine National Railways","Increasing The Burial Assistance For Veterans",
"Appropriating Funds for the Operation of the Government", "The Tax Incentives Management and Transparency Act (TIMTA)","Appropriating Funds for the Operation of the Government")
df4 <- df4[-14,]
rownames(df4) <- NULL
kable(df4,booktabs=T)%>% kable_styling(font_size = 5.5, full_width = F, latex_options = "hold_position") %>% footnote(general="All bills are enacted on a National Level")
```

Here are some of the significant laws that were passed under Robredo's run:

- RA10633: General Appropriations Act of 2014
- RA10638: Extending The Corporate Life Of The Philippine National Railways 
- RA10655: Decriminalizing Premature Marriages
- RA10661: National Children's Month Act
- RA10665: Open High School System Act
- RA10708: The Tax Incentives Management and Transparency Act (TIMTA)

---

## Authored Legislations

```{r table2, echo = FALSE, warning=FALSE}
options(knitr.kable.NA = '')

#Legislations authored by the candidate
authored_legislations <- legislations %>% filter(authorship == "primary author") %>% select(dates_filed, house_bill_no, full_title, short_title, significance, category, enacted_into_law, date_of_enactment)

df2 <- authored_legislations[,c("house_bill_no","short_title","significance")] 
colnames(df2) <- c("House Bill no.","Short Title", "Significance")
df2_1 <- df2[1:12,c(1,3)]
df2_2 <- df2[13:24,c(1,3)]
df2_3 <- df2[25:36,c(1,3)]
rownames(df2_1) <- NULL
rownames(df2_2) <- NULL
rownames(df2_3) <- NULL
kbl(list(df2_1, df2_2, df2_3),booktabs=T) %>% kable_styling(font_size = 6, full_width = F) %>% footnote(number="HB - House bill")
```

Notable house bills authored:

- HB00019: Full Disclosure Act Of 2013
- HB05452: Declaring The Second Week Of November Of Every Year As Economic And Financial Literacy Week
- HB06443: Strengthening The Barangays Through Support For Barangay Officials
- HB06467: Requiring Full Disclosure of Information on Fiscal Management 

---

## Co-Authored / Sponsored Legislations

```{r table3, echo = FALSE, warning=FALSE}
options(knitr.kable.NA = '')

#Legislations co-authored/sponsored by the candidate
co_authored_legislations <- legislations %>% filter(authorship == "co-author") %>% select(dates_filed, house_bill_no, full_title, short_title, significance, category, enacted_into_law, date_of_enactment)

df3 <- co_authored_legislations[,c("house_bill_no","short_title","significance")] 
colnames(df3) <- c("House Bill no.","Short Title", "Significance")
df3_1 <- df3[1:12,c(1,3)]
df3_2 <- df3[13:24,c(1,3)]
df3_3 <- df3[25:36,c(1,3)]
rownames(df3_1) <- NULL
rownames(df3_2) <- NULL
rownames(df3_3) <- NULL
kbl(list(df3_1, df3_2, df3_3),booktabs=T) %>% kable_styling(font_size = 6, full_width = F) %>% footnote(number="HB - House bill")
```

Notable house bills co-authored/sponsored:

- HB00050: Special Education Act
- HB00108: National Land Use And Management Act Of The Philippines
- HB00110: Anti-discrimination Act
- HB00191: Department Of Overseas Workers Act
- HB01214: Requiring The Use Of Filipino Sign Language Insets For Local News Programs

---

## Co-Authored / Sponsored Legislations (cont..)

```{r table3_1, echo = FALSE, warning=FALSE}
options(knitr.kable.NA = '')

#Legislations co-authored/sponsored by the candidate
df3_1 <- df3[37:48,c(1,3)]
df3_2 <- df3[49:60,c(1,3)]
df3_3 <- df3[61:72,c(1,3)]
rownames(df3_1) <- NULL
rownames(df3_2) <- NULL
rownames(df3_3) <- NULL
kbl(list(df3_1, df3_2, df3_3),booktabs=T) %>% kable_styling(font_size = 6, full_width = F) %>% footnote(number="HB - House bill")
```

- HB03587: The Anti - Political Dynasty Act Of 2013
- HB04113: People's Participation in the Budget Deliberation Act
- HB04382: Instituting A National Land Use And Management Policy
- HB04544: Revised Anti- Carnapping Act
- HB04558: Establishing Special Education (SPED) Centers For Children With Special Needs in all Public Schools Divisions

---

## Co-Authored / Sponsored Legislations (cont..)

```{r table3_2, echo = FALSE, warning=FALSE}
options(knitr.kable.NA = '')

#Legislations co-authored/sponsored by the candidate
df3_1 <- df3[73:84,c(1,3)]
df3_2 <- df3[85:96,c(1,3)]
rownames(df3_1) <- NULL
rownames(df3_2) <- NULL
kbl(list(df3_1, df3_2),booktabs=T) %>% kable_styling(font_size = 6, full_width = F) %>% footnote(number="HB - House bill")
```

- HB04688: Accreditation of NGOs and POs Eligible to Access Government Funds
- HB05411: Strengthening The Right Of The People To Free Expression
- HB05675: Defining Electronic Violence Against Women And Their Children 
- HB05801: Strengthening The Right Of Citizens To Information Held By The Government
- HB06135: Establishing The Coconut Farmers And Industry Trust Fund

---

## Summary of Findings

- Overall, Robredo performed better than at least 78% of her colleagues in the 16th Congress in terms of total bills filed and better than at least 90% of her colleagues in terms of number of days filed.

- Majority of Robredo's legislations are centered on economics (funds, management), checks and balances (government, national, system, penalties), and human welfare (children, sped, development, education)

- Majority of Robredo's legislations are intended to change the technicalities or reform existing laws (Rules and Revision of Laws) followed by categories under specific areas such as Welfare of Children, Appropriations, and Public Works and Highways 

- At least out of four bills were found to be authored mainly by Robredo.

- Most of Robredo's legislations were significant on a national level.

---

## Conclusion

The main objective of this presentation is to understand the legislative career of the Leni Robredo as a congresswoman. Based on the findings above, it is evident that the candidate is qualified to hold the position of the presidency due to the following reasons: (1) her performance has proven her to be a competent and productive politician compared to the rest of the members of the 16th Congress, (2) her authored and sponsored legislations are perfectly aligned with her background as an economics and law graduate as well as her personal beliefs, and (3) her significant contributions have demonstrated her willingness and competency to understand issues of national interest.